Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-07-22 analysis2_manuscript.tex
Clarify sweep conditioning

    % sweep interface
    While the \texttt{DFE} class facilitates the simulation of non-neutral variation across a sequence,
    it does not provide a means of specifying the origin and fate of particular selected alleles.
    Thus, we have also implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that enables 
    selective sweeps to be simulated with a high degree of control.
    This function augments a demographic model by introducing an advantageous mutation over a given time interval,
    at a given position along the sequence; and conditions the simulation such that the mutation is above a particular 
    frequency at the beginning and end of the sweep period.
    The conditioning is implemented via rejection sampling; if the frequency of the advantageous allele 
    drifts out of the prescribed bounds, the simulation is restarted from the introduction of
    % DRS: not editing this because I haven't looked at the implementation, but this doesn't seem right.
    % Obviously the allele must be allowed to drift outside of the bounds during the simulation, as long as it finishes
    % within those bounds. So if it is fixed during the sim when the bounds don't extend to 1.0 (or lost when the
    % bounds don't extent to 0.0) then we restart, and we restart if at the end of the sweep period we are outside of
    % the bounds. right? If what I am saying is correct then maybe this should be rephrased, and more importantly the
    % corresponding part of the methods should be corrected.
    the advantageous allele. If there are competing advantageous alleles (e.g. simultaneous sweeps at distinct loci) 
    then the simulation is restarted from the introduction of the first of these.
    As this functionality is implemented on top of the existing \stdpopsim API, it may be combined with other 
    models of selection and demography to simulate the combined effects of multiple, disparate evolutionary processes
    on genetic variation in a population.

@@

    Selective sweeps are implemented in the \slim engine by
    introducing a mutation with a fitness effect of $h s$ where $h$ is the
    dominance coefficient (as for the DFE model) and $s \geq 0$ is the
    selection coefficient.  The allele is introduced in a given population at
    a specified time and genomic position, and may be either globally or locally
    adaptive (so that the fitness effect is confined to the population of
    origin) over a fixed duration. To model sweeps from standing genetic
    variation (so-called ``soft sweeps''), the allele may be introduced at a
    given frequency, in which case carriers are selected at random at the
    initiation of the sweep.

    Crucially, the advantageous allele must remain extant over the course of
    the sweep; this conditioning is accomplished via rejection sampling,
    whereby the simulation is restarted from the onset of the sweep if the
    allele drifts out of the population of origin (or alternatively, below a
    user-specified frequency). If there are competing advantageous alleles
    % DRS: see my earlier comment about how we implement the conditioning here
    % and correct as needed
    (simultaneous sweeps at distinct loci) and one of these drifts out of
    bounds, then the simulation is restarted from the onset of the first sweep;
    this avoids conditioning on particular (partial) allele frequency
    trajectories for the remaining alleles.  A downside of this strategy is
    that rejection becomes more likely with increasing numbers of simultaneous
    sweeps or with hard-to-satisfy constraints; but it allows for greater
    flexibility in sweep specification and for interoperability with the
    \stdpopsim DFE model.

>>>>

    % sweep interface
    While the \texttt{DFE} class facilitates the simulation of non-neutral
    variation across a sequence, it does not provide a means of specifying the
    origin and fate of particular selected alleles.  Thus, we have also
    implemented an API in \stdpopsim (\texttt{stdpopsim.selective\_sweep}) that
    enables selective sweeps to be simulated with a high degree of control.
    This function augments a demographic model by introducing a mutation at a
    particular time and position along the sequence; sets the selection
    coefficient of the mutation to a positive value over a given time interval;
    and conditions the simulation such that the mutation is above a particular
    frequency at the beginning and end of the sweep period.  The conditioning
    is implemented via rejection sampling; if the advantageous allele is lost
    or does not have a frequency exceeding a user-specified threshold at the
    beginning or end of the sweep, the simulation is restarted from the
    introduction of the advantageous allele.  If there are competing
    advantageous alleles (e.g.  simultaneous sweeps at distinct loci) then the
    simulation is restarted from the introduction of the first of these.  As
    this functionality is implemented on top of the existing \stdpopsim API, it
    may be combined with other models of selection and demography to simulate
    the combined effects of multiple, disparate evolutionary processes on
    genetic variation in a population.

@@

    Selective sweeps are implemented in the \slim engine by introducing a
    mutation with a fitness effect of $h s$ where $h$ is the dominance
    coefficient (as for the DFE model) and $s \geq 0$ is the selection
    coefficient. The allele is introduced in a given population at a specified
    time and genomic position, and may be either globally or locally adaptive
    (so that the fitness effect is confined to the population of origin) over a
    fixed duration. The start of the sweep may be offset from the introduction
    of the allele, to model selection on standing genetic variation (so-called
    ``soft sweeps'').

    Crucially, the advantageous allele must remain extant over the course of
    the sweep; this conditioning is accomplished via rejection sampling,
    whereby the simulation is restarted from the onset of the sweep if the
    allele drifts out of the population of origin. The conditioning may
    optionally be extended to reject trajectories where the frequency at the
    beginning and end of the sweep period is below a specified threshold. If
    there are competing advantageous alleles (simultaneous sweeps at distinct
    loci) and one of these does not satisfy the acceptance conditions, then the
    simulation is restarted from the onset of the first sweep; this avoids
    conditioning on particular (partial) allele frequency trajectories for the
    remaining alleles.  A downside of this strategy is that rejection becomes
    more likely with increasing numbers of simultaneous sweeps or with
    hard-to-satisfy constraints; but it allows for greater flexibility in sweep
    specification and for interoperability with the \stdpopsim DFE model.
